
---
title: "Investment Report for Projects in `r params$country`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  country: Bangladesh
  year_start: '2016-07-01'
  year_end: '2017-06-30'
  fy: 2017
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE)
```

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(kableExtra)
investment_annual_summary<-read_csv("investment_annual_summary.csv")
investment_services_projects <- read_csv("investment_services_projects.csv")
```


## Datasets 

### Investment Annual Summary
The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018. Source of this data are IFC and copied from data camp.
```{r investment-annual-summary}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```
For each region, the highest investments were made during the 2013 or 2014 fiscal years

```{r investment-region-summary, fig.cap="Figure 1.2 Investment regionwise"}
investment_region_summary <- investment_annual_summary %>%
      group_by(region) %>%
      summarise(investment_amount =sum(dollars_in_millions))

ggplot(investment_region_summary ,aes(x=region ,y= investment_amount))+
  geom_bar(stat = "identity",fill="purple" ,alpha= 0.8)+
  theme_minimal() +
  xlab("Region")+ylab('Dollars in Millions')+
  theme(axis.text.x = element_text(angle = 45 , hjust = 1))
```




```{r tables}
kable(investment_region_summary, col.names = c("Region", "Dollars in Millions"), align = "cc", caption = "Table 1.1 The total investment summary for each region for the 2012 to 2018 fiscal years.")
```







### Investment Projects in `r params$country`
The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts. Projects that do not have an associated investment amount are excluded from the plot.

```{r country-investment-projects}
country_investment_projects <- investment_services_projects %>%
  filter(country == params$country) 

ggplot(country_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```




### Investment Projects in `r params$country` in `r params$fy`
The `investment_services_projects` dataset was filtered below to focus on information about each investment project from the `r params$fy` fiscal year, and is referred to as `country_annual_investment_projects`. Projects that do not have an associated investment amount are excluded from the plot.
```{r country-annual-investment-projects}
country_annual_investment_projects <- investment_services_projects %>%
  filter(country == params$country,
         date_disclosed >= params$year_start,
         date_disclosed <= params$year_end) 

ggplot(country_annual_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  ) 
```


