---
title: 'Workshop on Data Analysis with R '
author: "Zahid Asghar"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  pdf_document:
    toc: yes
    toc_depth: '3'
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
vignette: |
  %\VignetteIndexEntry{examples} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
subtitle: Friday and Saturday 6:00 PM to 8:00 PM (Online)
header-includes:
- \usepackage{titling}
- \pretitle{ \begin{center} \hspace*{-16cm} \includegraphics[width=1.5in,height=1.5in]{/Users/hp//Pictures/umt.png}\LARGE\\
  \vspace*{-3cm} }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## Making Sense of Data

## Why R

Reproducibility, RMarkdown, Presentation, Webpages...., pdf, html, word, Latex and much more

# What will you learn

Basic feature of this workshop will be to start exploring data using R along-with hands-on experience. All workshop and related data will be provided to all the participants once they will register for the workshop.

## By the end of Data Analysis with R, you'll understand:

-   How to use base R and tidyverse to carry out analysis on your own data
-   The core features of a data analysis workflow
-   Generating summary statistics and visualizing data

## And you'll be able to use:

-   Basic functions in the tidyverse to process raw data for typical data analysis questions
-   Logical expressions and indexing to ask specific questions of data
-   Functions to conduct common statistical tests

## Why this is for you

-   You have your own data sets and you want to analyze it
-   You think R is important but intimidated by R and dont know where to start with
-   You want to do good in data analytics
-   You want to get data insight for your intelligent decision making
-   You want to be at par in academia with the rest of the world

# Uploaded data and required R packages

Understanding datasets requires many hours/days or in some cases weeks.There are many commercially available softwares but open source community based softwares have now dominated and R is one of these. R makes data understanding process as easy as possible through the dplyr package. It is one of the easiest solution for code-based data analysis. We will learn in this training how to do it. I have already uploaded videos for downloading packages and relevant information. [Installing R and RStudio](https://youtu.be/ZNBZevfYgo00),

[R Packages, Library, Errors](https://youtu.be/SRSNdCZ_QnE),

[R Codes Basics](https://youtu.be/0IEfQdjc5mM)

I have discussed the [Gapminder dataset](https://cran.r-project.org/web/packages/gapminder/index.html) in my videos and we shall use it throughout this training. It's available through CRAN, so make sure to install it. Here's how to load in all required packages:

```{r, echo=FALSE, warning=FALSE,message=FALSE}
library(tidyverse)
#install.packages("gapminder")
library(gapminder)
#gapminder %>% filter(country=="Pakistan")
```

# Ouline

## Some Basic Commands for data loading and its features

-   Reading csv files, excel, Stata , SPSS ... data files
-   View
-   glimpse
-   structure
-   head, tail

## Components

-   Column Selection
-   Data Filtering
-   Sort data with arrange
-   Creating Derived Columns
-   Calculating Summary Statistics
-   Summarising data groupwise
-   Creating Tables

## How to merge files

-   Merging files

### Summarising data by groups

Average life expectancy is calculated by_grouping continents:

```{r, echo=FALSE}
gapminder %>% 
  group_by(continent) %>% 
  filter(year==2007) %>% 
  summarize(avglife=mean(lifeExp)) 

```

### if_else command alongwith mutate

### Percentiles

Understanding percentiles w.r.t lifeexpectancy and gdp per capita for 200 countries using gapminder data.

## Advanced Analysis

## Visualizing Data

Creating awesome graphs using themes of popular magzines like **The Economist**, **Financial Times** and many ohters.

```{r, echo=FALSE}
library(ggthemes )
library(knitr)
gapminder_2007 <- gapminder %>% filter(year == 2007)
ggplot(data = gapminder_2007) +
geom_point(mapping = aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
scale_x_log10()

```

```{r, echo=FALSE}
library(plotly)
library(gapminder)
g <- crosstalk::SharedData$new(gapminder, ~continent)
gg <- ggplot(g, aes(gdpPercap, lifeExp, color = continent, frame = year)) +
  geom_point(aes(size = pop, ids = country)) +
  geom_smooth(se = FALSE, method = "lm") +
  scale_x_log10()
ggplotly(gg) %>% 
  highlight("plotly_hover")

```


## Basic of Regression Analysis

After exploratory analysis, we shall discuss modeling of data in this section.

# Pre-requisite for workshop preparation

-   install R and RStudio
-   install tidyverse package [R Packages, Library, Errors](https://youtu.be/SRSNdCZ_QnE),
-   install gapminder package [R Packages, Library, Errors](https://youtu.be/SRSNdCZ_QnE),
